MachLLensFinder - I: A Machine Learning Strong Lensing Identification Pipeline

### 2016-12-16 ms.tex
added Nan's section on simulated images for training/testing set

\documentclass{emulateapj}
\usepackage{amsmath} 
\usepackage{apjfonts} 
\usepackage{amssymb} 
\usepackage{xspace}
\usepackage{hyperref} 
\usepackage{natbib} 
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{epsf}
\usepackage{subfigure}
\bibliographystyle{apj_ads}
\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\comment}[1]{{\color{red} #1}}
\newcommand{\change}[1]{{\color{blue} #1}}
\newcommand{\todo}[1]{{\bf\color{blue} #1}}
%%%%%%%%%% User defined symbol %%%%%%
\def\gsim{\gtrsim}
\def\lsim{\lesssim}    
\def\Msun{M_\odot}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%-------------------------------------------------%
\begin{abstract} 
  Gravitational lensing offers a direct probe of the underlying mass
  distribution of lensing systems, a window to the high redshift
  universe, and a geometric probe of cosmological models.  The advent
  of large scale surveys such as the Large Synoptic Sky Telescope and
  Euclid has prompted a need for automatic and efficient
  identification of strong lensing systems.  In this first paper of a
  series, we present LensFinderML, a strong lensing identification
  pipeline that will be publicly released as open source software.

Giant arcs trace... cosmology... mass distribution.  Advent of DES,
LSST, etc. need an efficient way to identify strong lens images.  We
present {\sc{MACHLLENSFINDER}} (MACHine Learning LENSFINDER), an open
source image processing and deep learning.  This first paper focuses
on image Processing Techniques for Giant Arc Identification Tests with
mock images from the Hubble Space Telescope and LSST.  Image
processing improves by....
\end{abstract}
%-------------------------------------------------%

@@

%-------------------------------------------------%
\section{Methodology}
\label{sec:methods}
%-------------------------------------------------%

@@

Figure~\ref{fig:ROC} shows the ROC curves for three different HOG
parameterizations for the HST-like dataset in red, green, and blue.
The parameters and corresponding AUC for each curve are labeled in the
legend.  The models for the HST-like dataset perform well, and have
AUCs very close to 1.  On the other hand, the model for our LSST-like
dataset does not perform well.  The dashed diagonal like corresponds
to a model that is as good as a random classifier, e.g. a coin-flip
for each test image.  The LSST-like ROC curve does marginally better,
and the $AUC=0.527$ is within the statistical noise of the sample.

We train the model corresponding to each curve with the logistic
regression regularization parameter, $C_\text{LogReg}$, that best
describes the training dataset without overfitting.  We discuss the
choice of regularization parameter in
Section~\ref{sec:regularizedfit}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:ROC}
\begin{center}
\includegraphics[width=\columnwidth]{figures/ROC_multipleHOG.png}
\caption{Red, green, blue: ROC curves for a models trained on the
  HST-like data with the Logistic Regression algorithm.  The magenta
  shows the ROC curve for the LSST-like data.  The corresponding
  parameters are labeled, N for number of orientations, ppc for
  pixels-per-cell, cpb for cells-per-block, and C for best
  regularization parameter.  }
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Figure~\ref{fig:regularization} shows the model performance as a
function of regularization parameter.  The left panel shows the
accuracy, or score, of the model as a function of regularization
parameter.  The right panel shows the AUC of the model as a function
of regularization parameter.  The solid and dotted lines respectively
correspond to the model performance on the test and training set.

For either characterization of the model, the model performance on the
training set increases and asymptotes with $C_\text{LogReg}$.  With
increasing model complexity, the model better fits the data set it has
been trained on.  This is analogous to fitting a seventh order
polynomial to seven data points, where the fitting function will go
through every point.

EDITING HERE

@@

\item

@@

>>>>

\documentclass{emulateapj}
\usepackage{amsmath} 
\usepackage{apjfonts} 
\usepackage{amssymb} 
\usepackage{xspace}
\usepackage{hyperref} 
\usepackage{natbib} 
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{epsf}
\usepackage{subfigure}
\bibliographystyle{apj_ads}
\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\comment}[1]{{\color{red} #1}}
\newcommand{\change}[1]{{\it\color{purple} #1}}
\newcommand{\todo}[1]{{\bf\color{blue} #1}}
%%%%%%%%%% User defined symbol %%%%%%
\def\gsim{\gtrsim}
\def\lsim{\lesssim}    
\def\Msun{M_\odot}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%-------------------------------------------------%
\begin{abstract} 
Gravitational lensing offers a direct probe of the underlying mass
distribution of lensing systems, a window to the high redshift
universe, and a geometric probe of cosmological models.  The advent of
large scale surveys such as the Large Synoptic Sky Telescope and
Euclid has prompted a need for automatic and efficient identification
of strong lensing systems.  We present \change{LensFinderML}, a strong
lensing identification pipeline that will be publicly released as open
source software.  In this first application, we employ a fast feature
extraction method, Histogram of Oriented Gradients (HOG), to capture
edge features that are characteristic of strong gravitational arcs.
We use logistic regression to train a supervised classifier model on
the HOG of \change{HST- and LSST}-like images.  Our best performing
models reach a completeness rate of \change{XX\%} and purity rate of
\change{XX\%} for an HST-like sample.  However, for one orbit of LSST,
our model cannot distinguish between strong lensing systems and
non-strong lensing systems.  In absence of the brightest cluster
galaxy in LSST-like data, our trained model reaches a completeness of
\change{XX\%} and purity of \change{XX\%}.  Our results show one the
most efficient and effective methods for automated strong lensing
identification in space-based surveys.  The results also emphasize the
need for further development and testing of additional tools to
exploit[wc?]  ground-based surveys.
\end{abstract}
%-------------------------------------------------%

@@

%-------------------------------------------------%
\section{Methodology}
\label{sec:methods}
%-------------------------------------------------%
\subsection{Creating Simulated Images for Training Set and Testing Set}

We use mock Hubble Space Telescope (HST) and Large Synoptic Sky
Telescope (LSST) images generated with PICS (Pipeline for Images of
Cosmological Strong lensing) (Li et al. 2016) and LensPOP (Thomas
Colletti 2013) as training set and testing set. In these images, the
lensed images include the images of lens galaxies, lensed images, and
galaxies on the line of sight; the unlensed images include the images
of lens galaxies and the galaxies on the line of sight.

To produce simulated lensed galaxies, we first set mass models of the
lenses to be the Singular Isothermal Ellipsoid (SIE):

\begin{equation}
\kappa = \frac{\theta_{\rm E}}{2}\frac{1}{\sqrt{x_1^2/q+x_2^2 q}}.
\end{equation}

Where, $\theta_{\rm E}$ is the Einstein Radius, $q$ is axis
ratio. Einstein radius can be calculated according to the redshift of
the lens, the redshift of the source, and the velocity dispersion of
the lens galaxy.
\begin{equation}
\theta_{\rm E} = 4\pi\left(\frac{\sigma_v}{c}\right)\frac{D_{ls}}{D_{s}}, 
\end{equation}
where, $c$ is the speed of light, $\sigma_v$ is the velocity
dispersion of the lens galaxy, $D_{ls}$ and $D_{s}$ are the angular
diameter distance from the source plane to the lens plane and from the
source plane to the observer separately.

To rotate the lenses with a random orientation angles, we adopt the
transformation below:

\begin{eqnarray}
\label{eq:rotation}
\begin{bmatrix}

    x_{1}\prime      \\
    x_{2}\prime 
\end{bmatrix}
= 
\begin{bmatrix}
    \cos \phi  &  -\sin \phi      \\
    \sin \phi  &  ~\cos \phi      
\end{bmatrix} 
\begin{bmatrix}
    x_1      \\
    x_2      
\end{bmatrix} .
\end{eqnarray}

$\phi$ is the orientation angle.  According to above equations, we
need $\sigma_v$, $q$, $\phi$, redshift of the lens $z_l$, and redshift
of the source $z_s$ to derivative the lensing maps completely.
Velocity dispersion, ellipticity, and orientation are randomly chosen
from typical ranges of a galaxy, where $\sigma_v = [200, 320] km/s$,
$q = [0.5, 1.0]$, and $\phi = [0, 360]$. The redshift of the lens is
fixed at $z_l = 0.2$, and the redshift of the source is fixed at $z_s
= 1.0$. The images of source galaxies are extracted from the CANDELS
survey. We selected 33 bright galaxies, and relocated them to
projected positions, which are close to the caustic of the lensing
system, and the projected position of the lens is fixed in the center
of the field of view of each image. Once we have the lens, source, and
the parameters of the image simulation, i.e., $300 \times 300 {\rm
  pixels}^2$ and 0.03 arcsec per pixel per side, settled, the lensed
images can be produced by performing ray-tracing simulations.

The light distributions of the lens galaxies  are modeled as the elliptical Sersic profile,

\begin{equation}
I(R) = I_{\rm eff}~{\rm exp} \left\lbrace -b_{n} \left[ \left(
  \frac{R}{R_{\rm eff}}\right)^{1/n} - 1 \right ] \right\rbrace
\end{equation}

Where, $R = \sqrt[]{x_1^2 /q+x_2^2 q }$, $R_{\rm eff}$ is the
effective radius in units of arcsecond, $I_{\rm eff}$ is the intensity
at the effective radius, $n$ is the index of the Sersic profile, $q$
is the ellipticity, and We perform the similar transformation as
Eq. \ref{eq:rotation} to orient the sources.  We assume that the
distribution of light is following that of mass, so ellipticity and
orientation are the same as the SIE model. Matching the velocity
dispersion with the COSMOS morphological
catalog \footnote{\url{https://arxiv.org/pdf/1501.04977v2.pdf}}, we
assign effective radius, effective luminosity, and index to the light
profile.  We also assume the light center is on top of the mass
center, the noiseless images of lenses can be drawn. The galaxies on
the line of sight are constructed by cutting light-cones from the
Hubble Ultra Deep Field.  Stacking all these images together and
collaborating the magnitude of each component, we create primarily
simulated lensed images.

The final part is mock observing. To produce HST-like images, we do
not need to do anything further than stacking and magnitude
calibrating, because noise and PSF are involved when we are extracting
galaxies on the line of sight. To produce LSST-like images, we perform
re-binning, convolving with PSF, and adding noise serially; this
process is implemented in LensPop. We created single-epoch-LSST-like
images and stacked -LSST-like images for investigating the performance
and usability of our pipeline in LSST data. Figure 2.1 illustrates a
sample mock lensed and unlensed images from each telescope. From left
to right, the top panel images respectively correspond to a mock HST
lensing and non-lensing system. The mock HST images are 300 × 300 in
size, and the pixel size is 0.03 arcsecond. For the HST-like data-set,
many giant arcs are visually obvious. The middle panel images
respectively correspond to a mock single-epoch LSST lensing and
non-lensing systems. The bottom panel images respectively correspond
to a mock stacked LSST lensing and non-lensing systems. These images
are 45 × 45; the pixel size is 0.2 arcsecond. The ground based noise,
PSF, and limited resolution of the LSST-like data-set make visual
giant arc identification difficult. But when in the stacked LSST
images, the quality is better, because more data improves the signal
to noise ratio better.

In the training sets, we also generate unlensed images, the procedure
is similar to mock lensed images but getting rid of the part of
ray-tracing. 10000 lensed images and 10000 unlensed images are merged
to construct one training set. Furthermore, for investigating the
influence of BCG galaxies on the performance of our lens
identification pipeline in LSST data, we build another training sets
of the images without BCG galaxies.

@@

% HST model performance
Figure~\ref{fig:ROC} shows the ROC curves for three different HOG
parameterizations for the HST-like dataset in red, green, and blue.
The parameters and corresponding AUC for each curve are labeled in the
legend.  The models for the HST-like dataset perform well, and have
AUCs very close to 1.

% LSST model performance
On the other hand, the model for our LSST-like dataset (solid magenta)
does not perform well.  The dashed diagonal like corresponds to a
model that is as good as a random classifier, e.g. a coin-flip, for
each test image.  The LSST-like ROC curve does marginally better, and
the $AUC=0.527$ is within the statistical noise of the sample.

% No BCG LSST
One possible method of improving the detection of strong lensing
systems is to model out the BCG.  The dotted magenta ROC curve
corresponds to a model trained and tested on LSST-like data where the
BCG has not been added to the mock image.  This model results in an
AUC of 0.64, performing better than random.  In other words, if we
were to perfectly model out the BCG contribution in LSST-like data, we
could improve the performance by $\sim10\%$.

% Choosing C
The ROCs correspond to models that have been trained with logistic
regression regularization parameter, $C_\text{LogReg}$, that best
describes the training dataset without overfitting.  We discuss the
choice of regularization parameter in
Section~\ref{sec:regularizedfit}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:ROCHST}
\begin{center}
\includegraphics[width=\columnwidth]{figures/ROC_HST_HOG.png}
\caption{Red, green, blue: ROC curves for a models trained on the
  HST-like data with the Logistic Regression algorithm.  The magenta
  solid (dotted) curve show the ROC curve for the LSST-like data with
  (without) the BCG.  The corresponding parameters are labeled: N for
  number of orientations, ppc for pixels-per-cell, cpb for
  cells-per-block, and C for best regularization parameter.  The AUC
  for each model is also labeled in the legend, where an AUC close to
  1 is a near-ideal model, and an AUC close to 0.5 is a model that
  performs as well as a coin-flip.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:ROCLSST}
\begin{center}
\includegraphics[width=\columnwidth]{figures/ROC_LSST_HOG.png}
\caption{Red, green, blue: ROC curves for a models trained on the
  HST-like data with the Logistic Regression algorithm.  The magenta
  solid (dotted) curve show the ROC curve for the LSST-like data with
  (without) the BCG.  The corresponding parameters are labeled: N for
  number of orientations, ppc for pixels-per-cell, cpb for
  cells-per-block, and C for best regularization parameter.  The AUC
  for each model is also labeled in the legend, where an AUC close to
  1 is a near-ideal model, and an AUC close to 0.5 is a model that
  performs as well as a coin-flip.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%  Our running over C - score and AUC as characterizations
Figure~\ref{fig:regularization} shows the model performance as a
function of regularization parameter.  The left panel shows the
accuracy, or score, of the model as a function of regularization
parameter.  The accuracy is simply the percentage of correct
predictions, true positives and true negatives, out of a given set.
The right panel shows the AUC of the model as a function of
regularization parameter.  The solid and dotted lines respectively
correspond to the model performance on the test and training set.

% Training behavior with C
For either characterization of the model, the model performance on the
training set increases and asymptotes with $C_\text{LogReg}$.  With
increasing model complexity, the model better fits the data set it has
been trained on.  [wc???] This is analogous to fitting a seventh order
polynomial to seven data points, where the fitting function will go
through every point but will not likely predict additional points.

% Test behavior with C
We use the model performance on a test set to determine the best value
for $C_\text{LogReg}$ that will not overfit the training data.  With
increasing model complexity, we are able to better able to capture
features that are generally characteristic of strong lensing systems
with arcs.  However, past a certain $C_\text{LogReg}$, the model
performance on the test set decreases, as it has overfit the training
set.

@@

\item Publicly released pipeline will be... swap out feature
  extraction, machine learning algorithm, perform parameter searches,
  and train a model with best parameters.

\item HOG efficiently extracts features that well-describe strong
  lensing systems with an arc. Performance of the model... SVM doesn't
  make a difference.

\item Ground based... noise... Model out BCG.

@@

% Caveat of the mock data set
One major caveat ... training and test mock images... LOS galaxies all
from Hubble Ultra Deep field.  Success measured by performance on the
small sample of systems observed with HST.

%
